PANCHANGE_data_preprocessing

### 2020-08-25 PANCHANGE_analysis.Rmd
fixing age categories

```{r}
dat.raw %>%
  freq(age_category_full)
```

@@

@@

ocir.sum_score_cols
```

@@

+++ADD CHUNK FOR AGE CATEGORIES HERE
```{r Age categories}
dat %>%
  group_by(Sample) %>%
  count(age_category_full, Sample)
```

@@

@@

```{r gad percentage plot retro_prepan}
dat.long.gad.diff %>%
  filter(Time_point != "retro_prepan") %>%
  ggplot(mapping = aes(
    x = gad.diff_score,
    y = stat(density*width),
    fill = Time_point)) +
  geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Difference GAD7 score",
    y = "Percentage of participants",
    fill = "Comparison:\nBaseline vs") +
  theme_personal +
  scale_fill_manual(values = palette, labels = c("Prepandemic", "Retrospective"))
```

@@

@@

@@

@@

@@

@@

@@

>>>>

@@

## age_category_full
```{r Inspect age_category_full}
dat.raw %>%
  freq(age_category_full)

dat.raw %>%
  freq(age_category_full_numeric)
```

```{r Collapse age_category_full}
dat.raw <- dat.raw %>%
  mutate(age_category_collapsed_numeric =
           case_when(
             age_category_full_numeric == 1 ~ 0,
             age_category_full_numeric == 2 ~ 1,
             age_category_full_numeric == 3 ~ 2,
             age_category_full_numeric == 4 ~ 3,
             age_category_full_numeric == 5 ~ 4,
             age_category_full_numeric == 6 ~ 5,
             age_category_full_numeric == 7 ~ 6,
             age_category_full_numeric == 8 ~ 7,
             age_category_full_numeric == 9 ~ 9,
             age_category_full_numeric == 10 ~ 9,
             age_category_full_numeric == 11 ~ 9,
             age_category_full_numeric == 12 ~ 9,
             age_category_full_numeric == 13 ~ 9

           )
  )

dat.raw <- dat.raw %>%
  mutate(age_category_collapsed =
           recode_factor(
             age_category_collapsed_numeric,
             `0` = "16-18",
             `1` = "19-25",
             `2` = "26-35",
             `3` = "36-45",
             `4` = "46-55",
             `5` = "56-65",
             `6` = "66-70",
             `7` = "71-75",
             `8` = "76+"
           )
  )

dat.raw %>%
  freq(age_category_collapsed)
```

@@

ocir.sum_score_cols

pcl.sum_score_cols <- dat %>%
  select(contains("pcl")) %>%
  select(contains("sum_score")) %>% 
  select(-ends_with("raw")) %>%
  colnames()

pcl.sum_score_cols
```

@@

+++ADD CHUNK FOR AGE CATEGORIES HERE
```{r Age categories by Sample}
dat %>%
  group_by(Sample) %>%
  count(age_category_full, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))
```

@@

```{r gad gather to long format Time_points gad.sum_score}
dat.long.gad.sum <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "gad.sum_score",
    all_of(gad.sum_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Disorder_hierarchical,
    Time_point_raw,
    gad.sum_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "gad.sum_score_", "")) %>%
  select(-Time_point_raw)

head(dat.long.gad.sum)
```

### Plots diff

@@

```{r gad percentage plot retro_prepan diff}
dat.long.gad.diff %>%
  filter(Time_point != "retro_prepan") %>%
  ggplot(mapping = aes(
    x = gad.diff_score,
    y = stat(density*width),
    fill = Time_point)) +
  geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Difference GAD7 score",
    y = "Percentage of participants",
    fill = "Comparison:\nBaseline vs") +
  theme_personal +
  scale_fill_manual(values = palette, labels = c("Prepandemic", "Retrospective"))
```

@@

### Plots sum

```{r gad percentage plot retro_prepan sum}
dat.long.gad.sum %>%
#  filter(Time_point != "retro_prepan") %>%
  ggplot(mapping = aes(
    x = gad.sum_score,
    y = stat(density*width),
    fill = Time_point)) +
  geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "GAD7 scores",
    y = "Percentage of participants",
    fill = "Time point") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette3,
    labels = c("Baseline", "Prepandemic", "Retrospective")
    )
```

```{r gad percentage plot retro_prepan sum, fig.height=6, fig.width=10}
dat.long.gad.sum %>%
#  filter(Time_point != "retro_prepan") %>%
  ggplot(mapping = aes(
    x = gad.sum_score,
    y = stat(density*width),
    fill = Time_point)) +
  geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "GAD7 score",
    y = "Percentage of participants",
    fill = "Time point") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette3,
#    labels = c("Prepandemic", "Retrospective")
    ) +
  facet_wrap("Disorder_hierarchical")
```

@@

```{r phq gather to long format Time_points phq.sum_score}
dat.long.phq.sum <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "phq.sum_score",
    all_of(phq.sum_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Disorder_hierarchical,
    Time_point_raw,
    phq.sum_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "phq.sum_score_", "")) %>%
  select(-Time_point_raw)

head(dat.long.phq.sum)
```

@@

### Plots sum

```{r phq percentage plot retro_prepan sum}
dat.long.phq.sum %>%
  #  filter(Time_point != "retro_prepan") %>%
  ggplot(mapping = aes(
    x = phq.sum_score,
    y = stat(density*width),
    fill = Time_point)) +
  geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "PHQ9 scores",
    y = "Percentage of participants",
    fill = "Time point") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette3,
    labels = c("Baseline", "Prepandemic", "Retrospective")
  )
```

```{r phq percentage plot retro_prepan sum, fig.height=6, fig.width=10}
dat.long.phq.sum %>%
  #  filter(Time_point != "retro_prepan") %>%
  ggplot(mapping = aes(
    x = phq.sum_score,
    y = stat(density*width),
    fill = Time_point)) +
  geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "PHQ9 score",
    y = "Percentage of participants",
    fill = "Time point") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette3,
    labels = c("Baseline", "Prepandemic", "Retrospective")
  ) +
  facet_wrap("Disorder_hierarchical")
```

@@

```{r ocir gather to long format Time_points ocir.sum_score}
dat.long.ocir.sum <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "ocir.sum_score",
    all_of(ocir.sum_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Disorder_hierarchical,
    Time_point_raw,
    ocir.sum_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "ocir.sum_score_", "")) %>%
  select(-Time_point_raw)

head(dat.long.ocir.sum)
```

@@

### Plots sum

```{r ocir percentage plot retro_prepan sum}
dat.long.ocir.sum %>%
  ggplot(mapping = aes(
    x = ocir.sum_score,
    y = stat(density*width),
    fill = Time_point)) +
  geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "OCI-R scores",
    y = "Percentage of participants",
    fill = "Time point") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette3,
    labels = c("Baseline", "Prepandemic", "Retrospective")
  )
```

```{r ocir percentage plot retro_prepan sum, fig.height=6, fig.width=10}
dat.long.ocir.sum %>%
  ggplot(mapping = aes(
    x = ocir.sum_score,
    y = stat(density*width),
    fill = Time_point)) +
  geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "OCI-R score",
    y = "Percentage of participants",
    fill = "Time point") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette3,
    labels = c("Baseline", "Prepandemic", "Retrospective")
  ) +
  facet_wrap("Disorder_hierarchical")
```

@@

```{r pcl gather to long format Time_points pcl.sum_score}
dat.long.pcl.sum <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "pcl.sum_score",
    all_of(pcl.sum_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Disorder_hierarchical,
    Time_point_raw,
    pcl.sum_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "pcl.sum_score_", "")) %>%
  select(-Time_point_raw)

head(dat.long.pcl.sum)
```

@@

### Plots sum

```{r pcl percentage plot retro_prepan sum}
dat.long.pcl.sum %>%
  ggplot(mapping = aes(
    x = pcl.sum_score,
    y = stat(density*width),
    fill = Time_point)) +
  geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "PCL scores",
    y = "Percentage of participants",
    fill = "Time point") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette3,
    labels = c("Baseline", "Prepandemic", "Retrospective")
  )
```

```{r pcl percentage plot retro_prepan sum, fig.height=6, fig.width=10}
dat.long.pcl.sum %>%
  ggplot(mapping = aes(
    x = pcl.sum_score,
    y = stat(density*width),
    fill = Time_point)) +
  geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "PCL score",
    y = "Percentage of participants",
    fill = "Time point") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette3,
    labels = c("Baseline", "Prepandemic", "Retrospective")
  ) +
  facet_wrap("Disorder_hierarchical")
```

### 2020-08-25 PANCHANGE_data_preprocessing2020-07-15.Rmd
fixing age categories

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(`Login ID`) %>% # Drop NAs
    distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
#      startDate.coping, #Date when participant signed up to ramp  (is in the coping gad chunk)
      age_category_full = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
#     EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
#     demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
#     demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#     demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#     demographics.cses_or_equivalent, # level of education - cse
#     demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity_unc = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

dem.raw.id <- dem.raw.id %>%
    mutate(
      age_category_full =
        recode_factor(
          age_category_numeric, 
          "1" = "16-18",
          "2" = "19-25",
          "3" = "26-35",
          "4" = "36-45",
          "5" = "46-55",
          "6" = "56-65",
          "7" = "66-70",
          "8" = "71-75",
          "9" = "76-80",
          "10" = "81-85",
          "11" = "86-90",
          "12" = "91-100",
          "13" = "100+"
          )
    )

dem.raw.id %>%
  freq(age_category_full)
}
```

@@

```{r Exclude participants that do not report age}
dim(dat.no.sex.no.gender)
dat.no.sex.no.gender.no.age <- dat.no.sex.no.gender %>%
  drop_na(age_category_full)
dim(dat.no.sex.no.gender.no.age)

@@

Age is not available in RAMP, only categorical age. We will assign each individual the middle value of each age category. 
```{r create age - proxy variable for ramp}
if(RAMP == TRUE){
dat.coping <- dat.coping %>%
  mutate(
    Age_respiratory =
      case_when(
        age_category_full == "16-18" ~ 17,
        age_category_full == "19-25" ~ 22,
        age_category_full == "26-35" ~ 30.5,
        age_category_full == "36-45" ~ 40.5,
        age_category_full == "46-55" ~ 50.5,
        age_category_full == "56-65" ~ 60.5,
        age_category_full == "66-70" ~ 68,
        age_category_full == "71-75" ~ 73,
        age_category_full == "76-80" ~ 78,
        age_category_full == "81-85" ~ 83,
        age_category_full == "86-90" ~ 88,
        age_category_full == "91-100" ~ 93,
      age_category_full == "100+" ~ 100
      )
  )
}

>>>>

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(`Login ID`) %>% # Drop NAs
    distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
#      startDate.coping, #Date when participant signed up to ramp  (is in the coping gad chunk)
      age_category = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
#     EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
#     demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
#     demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#     demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#     demographics.cses_or_equivalent, # level of education - cse
#     demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity_unc = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

dem.raw.id <- dem.raw.id %>%
    mutate(
      age_category =
        recode_factor(
          age_category_numeric, 
          "1" = "16-18",
          "2" = "19-25",
          "3" = "26-35",
          "4" = "36-45",
          "5" = "46-55",
          "6" = "56-65",
          "7" = "66-70",
          "8" = "71-75",
          "9" = "76-80",
          "10" = "81-85",
          "11" = "86-90",
          "12" = "91-100",
          "13" = "100+"
          )
    )

dem.raw.id %>%
  freq(age_category)
}
```

@@

```{r Exclude participants that do not report age}
dim(dat.no.sex.no.gender)
dat.no.sex.no.gender.no.age <- dat.no.sex.no.gender %>%
  drop_na(age_category)
dim(dat.no.sex.no.gender.no.age)

@@

Age is not available in RAMP, only categorical age. We will assign each individual the middle value of each age category. 
```{r create age - proxy variable for ramp}
if(RAMP == TRUE){
dat.coping <- dat.coping %>%
  mutate(
    Age_respiratory =
      case_when(
        age_category == "16-18" ~ 17,
        age_category == "19-25" ~ 22,
        age_category == "26-35" ~ 30.5,
        age_category == "36-45" ~ 40.5,
        age_category == "46-55" ~ 50.5,
        age_category == "56-65" ~ 60.5,
        age_category == "66-70" ~ 68,
        age_category == "71-75" ~ 73,
        age_category == "76-80" ~ 78,
        age_category == "81-85" ~ 83,
        age_category == "86-90" ~ 88,
        age_category == "91-100" ~ 93,
      age_category == "100+" ~ 100
      )
  )
}
